---
title: "NHS Workforce Statistics for `r params$ics_name`"
subtitle: "Data for `r params$month_year`"
author: "Maria Garcia"
date: "2025-12-03"
params:
  ics_name: "North East and North Cumbria"
  month_year: "July 2025"
format:
  html:
    toc: true
    toc-location: left
    toc-title: Contents
execute:
  echo: false
  message: false
  warning: false
---

```{r load libraries}
library("bsicons")
library("bslib")
library("dplyr")
library("ggplot2")
library("glue")
library("reactable")
library("readxl")
library("tidyr")
```

```{r, read in data}
filename = glue("NHS Workforce Statistics, {params$month_year} England and Organisation.xlsx")
staff_group = read_xlsx(path = filename, sheet = "3. NHSE, Org & SG - FTE", skip = 5)
time_series = read_xlsx(path = filename, sheet = "5. All Staff, NHSE & Org - FTE", skip = 4)
```


```{r tidy staff_group}
staff_group = 
  staff_group |>
  fill(`ICS name`) |>
  filter(`ICS name` == params$ics_name) |>
  select(`Organisation name`, `Total`,
         `HCHS Doctors`, `Nurses & health visitors`,
         `Midwives`, `Ambulance staff`) |>
  replace_na(list(`Organisation name` = "All"))
```

## Summary

This report shows monthly numbers of NHS Hospital and Community Health Services (HCHS) staff working in NHS Trusts and other core organisations in `r params$ics_name`. 

## Key facts

```{r value boxes}
total_fte = 
  staff_group |>
  filter(`Organisation name` == "All") |>
  pull(Total) |> 
  round() |>
  prettyNum(big.mark = ",")

value_box(
  title = "All NHS Hospital & Community Health Service (HCHS) staff",
  value = glue("{total_fte} FTE in {params$month_year}"),
  showcase = bs_icon("hospital"),
  theme = "primary"
)
```

### Information `r bs_icon("info-circle")`
*Full-Time Equivalent (FTE)* is a standardised measure of the workload of an employed person and allows for the total workforce workload to be expressed in an equivalent number of full-time staff. 1.0 FTE equates to full-time work of 37.5 hours per week.

## Trends

The graph below presents time series data for the three biggest NHS organisations in `r params$ics_name` from January 2022 to `r params$month_year`.

```{r tidy time_series}
time_series = 
  time_series |>
  fill(`ICS name`) |>
  filter(`ICS name` == params$ics_name) |>
  select(-(`NHS England region code`:`ICS name`), -`Organisation code`) |>
  pivot_longer(-`Organisation name`, names_to = "date", values_to = "count") |>
  mutate(date = as.Date(as.numeric(date), origin = "1899-12-30")) |> 
  filter(!is.na(`Organisation name`))
```

```{r plot}
# Three biggest organisations

biggest_orgs = 
  time_series |> 
  filter(date == max(date)) |>
  slice_max(n = 3, `count`) |>
  pull(`Organisation name`)

plot = 
  time_series |>
  filter(`Organisation name` %in% biggest_orgs) |>
  filter(date > "2021-12-01") |>
  ggplot(aes(x = date, y = count, colour = `Organisation name`)) + 
  geom_point() + 
  geom_line() +
  theme_minimal() + 
  xlab("FTE staff") +
  ylim("")

plotly::ggplotly(plot)
```

## Table

```{r most_midwives}
max_midwives = 
  staff_group |>
  filter(`Organisation name` != "All") |>
  slice_max(`Midwives`, n = 1) 
```

The table below shows the total number of doctors and nurses for each organisation within `r params$ics_name`. We can see that the organisation with the most midwives is the _`r pull(max_midwives, "Organisation name")`_ with `r round(max_midwives$Midwives)` staff.

```{r table}
staff_group |>
  reactable(style = list(fontSize = "0.8rem"),
            defaultColDef = colDef(
              format = colFormat(digits = 0)
            ),
            columns = list(
              `Organisation name` = colDef(minWidth = 140)  
            ),
            bordered = TRUE,
            highlight = TRUE) 
```

